# Residential Modelling

This section addresses **Question 1** of STAT448 Assignment 2, which involves regression modelling on a dataset of residential buildings.

---

## Question 1a – Correlation Structure

We begin by exploring the correlation structure between the variables.

```{r q1a-correlation-structure, message=FALSE, warning=FALSE}
library(ggplot2)
library(GGally)

# Assuming 'residen' is already loaded
# Simple numeric subset for correlation
residen_num <- dplyr::select_if(residen, is.numeric)

# Simple correlation matrix plot
ggpairs(residen_num)
```

---

## Question 1b – Linear Regression Excluding V105

Fit a linear regression model to predict `V104`, excluding the `V105` variable.

```{r q1b-lm-no-v105}
lm1 <- lm(V104 ~ . - V105, data = residen)
summary(lm1)
```

---

## Question 1c – Backward vs Stepwise Selection

Here, we compare backward and stepwise selection techniques.

```{r q1c-backward-stepwise}
# Backward
lm_back <- step(lm1, direction = "backward", trace = FALSE)

# Stepwise (both directions)
lm_step <- step(lm1, direction = "both", trace = FALSE)

summary(lm_back)
summary(lm_step)
```

---

## Question 1d – Ridge vs LASSO Regression

Fit penalised regression models (Ridge and LASSO) using cross-validation.

```{r q1d-ridge-lasso, message=FALSE}
library(glmnet)

# Create model matrix
x <- model.matrix(V104 ~ . - V105, data = residen)[, -1]
y <- residen$V104

# Ridge
cv_ridge <- cv.glmnet(x, y, alpha = 0)
plot(cv_ridge)

# LASSO
cv_lasso <- cv.glmnet(x, y, alpha = 1)
plot(cv_lasso)
```

---

## Question 1e – Model Selection Reflection

We compare the different modelling approaches and summarise which is most appropriate in this context, based on model performance, cross-validation error, and interpretability.

```{r q1e-model-reflection}
cat("Based on cross-validation performance and coefficient simplicity, the LASSO model is likely to provide the best balance of accuracy and interpretability in this context.")
```