```{r setup, include=FALSE}
# This chunk was set up with the aid of ChatGPT.
# The intent is to load updates quietly thus not
# spending undue time with the logistics of getting 
# setup. 

#---------------- loading libraries ---------------------------
#---------------- loading libraries ---------------------------
# Load configuration for reproducibility and preferred CRAN mirror
options(repos = c(CRAN = "https://cran.stat.auckland.ac.nz/"))

# Required packages
required_packages <- c(
  "conflicted",    # Conflict resolution
  "ggplot2",       # Core plotting
  "dplyr",         # Data wrangling
  "tidyverse",     # Core data science packages
  "flextable",     # Summary tables
  "skimr",         # Data summaries
  "GGally",        # Pair plots
  "MASS",          # LDA/QDA and logistic
  "class",         # kNN
  "caret",         # Model training utilities
  "glmnet",        # Regularised regression
  "cowplot",       # Plot composition
  "knitr",         # Inline rendering
  "kableExtra"     # Table formatting
)

for (pkg in required_packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg)
  }
  library(pkg, character.only = TRUE)
}

#---------------- conflict_prefer ------------------------------
#---------------- conflict_prefer ------------------------------
conflict_prefer("filter", "dplyr"); conflict_prefer("select", "dplyr")

set.seed(82171165)
knitr::opts_chunk$set(dev.args = list(png = list(type = "cairo")))


# ---- Pretty Summary  with Type Summary and Preview ------
# ---- Pretty Summary  with Type Summary and Preview ------
# works with flextable; provides a summary of column types
# and a preview

pretty_summary <- \(df) {   # works with flextable 
  select(df, where(is.numeric)) %>%
    summarise(
      across(
        everything(),
        .fns = list(
          Mean     = \(x) mean(x, na.rm = TRUE),
          Median   = \(x) median(x, na.rm = TRUE),
          Min      = \(x) min(x, na.rm = TRUE),
          Max      = \(x) max(x, na.rm = TRUE),
          IQR      = \(x) IQR(x, na.rm = TRUE),
          nNA      = \(x) sum(is.na(x))
        )
      )
    ) %>%
    pivot_longer(
      cols = everything(),
      names_to = c("Variable", ".value"),
      names_sep = "_"
    ) %>%
    mutate(across(where(is.numeric), round, 2)) %>%
    flextable() %>%
    set_header_labels(
      Variable = "Feature",
      Mean     = "Mean",
      Median   = "Median",
      Min      = "Min",
      Max      = "Max",
      IQR      = "Interquartile Range",
      nNA      = "Missing Values"
    ) %>%
    autofit() %>%
    theme_vanilla()
}

# ---- Pretty csv_read with Type Summary and Preview ------
# ---- Pretty csv_read with Type Summary and Preview ------ 
# works with flextable; provides a summary of column types
# and a preview

pretty_read_csv <- \(path, n = 0) {                   # to work with flextable
  df <- readr::read_csv(path, show_col_types = FALSE)
  types <- purrr::map_chr(df, typeof)
  type_df <- data.frame(Column = names(df), Type = types, stringsAsFactors = FALSE)

  print(flextable::flextable(type_df) |>
          flextable::autofit() |>
          flextable::set_caption(caption = "Column Types in Loaded CSV"))

  if (n > 0) print(head(df, n))
  return(df)
}

# ---- Pretty Excel Reader with Type Summary and Preview ------
# Uses flextable for consistent formatting with pretty_read_csv()

pretty_read_xlsx <- \(path, sheet = 1, n = 0) {
  # Read Excel file (defaults to first sheet)
  df <- readxl::read_excel(path, sheet = sheet)
  
  # Summarise column types
  types <- purrr::map_chr(df, typeof)
  type_df <- data.frame(Column = names(df), Type = types, stringsAsFactors = FALSE)
  
  # Display types in a flextable
  print(flextable::flextable(type_df) |>
          flextable::autofit() |>
          flextable::set_caption(caption = "Column Types in Loaded XLSX"))
  
  # Optionally preview first n rows
  if (n > 0) print(head(df, n))
  
  return(df)
}


pretty_df <- function(df, title = NULL, fontsize = 10, max_rows = 20) {
  df_show <- head(df, max_rows)  # Limit to first `max_rows` rows for readability
  
  ft <- flextable(df_show) %>%
    fontsize(size = fontsize, part = "all") %>%
    autofit() %>%
    theme_booktabs() %>%
    align(align = "center", part = "all") %>%
    padding(padding = 5, part = "all") 
  
  if (!is.null(title)) {
    ft <- set_caption(ft, caption = title)
  }
  
  return(ft)
}

library(flextable)
library(ggplot2)

pretty_ggplot <- function(plot, title = "ggplot Summary") {
  if (!inherits(plot, "gg")) stop("Input must be a ggplot object.")

  # Extract key components
  plot_data <- tryCatch(plot$data, error = function(e) NULL)
  geoms <- sapply(plot$layers, function(layer) class(layer$geom)[1])
  mappings <- plot$mapping
  
  # Pull out global aesthetics as strings
  global_aes <- names(mappings)
  global_vals <- sapply(mappings, function(x) rlang::expr_text(x))

  # Additional metadata
  plot_title <- plot$labels$title %||% ""
  x_lab <- plot$labels$x %||% ""
  y_lab <- plot$labels$y %||% ""
  colour_lab <- plot$labels$colour %||% plot$labels$color %||% ""

  # Build a metadata data frame
  df <- data.frame(
    Component = c("Title", "X Axis", "Y Axis", "Colour Legend", "Geoms", global_aes),
    Value = c(plot_title, x_lab, y_lab, colour_lab, paste(geoms, collapse = ", "), global_vals),
    stringsAsFactors = FALSE
  )

  # Format with flextable
  flextable(df) %>%
    set_caption(caption = title) %>%
    autofit() %>%
    theme_booktabs() %>%
    fontsize(size = 10, part = "all") %>%
    align(align = "left", part = "all") %>%
    padding(padding = 4)
}

library(flextable)
library(broom)

library(flextable)
library(broom)

pretty_model <- function(model, title = "Model Summary", digits = 3) {
  if (!inherits(model, "glm")) stop("Only 'glm' objects are currently supported.")
  
  # Use broom::tidy() for clean coefficient table
  tidy_tbl <- broom::tidy(model) %>%
    mutate(across(where(is.numeric), ~ round(.x, digits)))

  # Optionally include model fit stats (like AIC, null deviance, etc.)
  stats_tbl <- data.frame(
    Metric = c("Formula", "AIC", "Null deviance", "Residual deviance"),
    Value = c(
      as.character(formula(model)),
      round(AIC(model), digits),
      round(model$null.deviance, digits),
      round(model$deviance, digits)
    ),
    stringsAsFactors = FALSE
  )
  
  # Format coefficient table
  ft_coef <- flextable(tidy_tbl) %>%
    set_caption(caption = title) %>%
    autofit() %>%
    theme_booktabs() %>%
    fontsize(size = 10, part = "all") %>%
    align(align = "center", part = "all") %>%
    padding(padding = 4, part = "all")

  # Format stats table separately (optional)
  ft_stats <- flextable(stats_tbl) %>%
    set_caption(caption = paste(title, "(Fit Stats)")) %>%
    autofit() %>%
    theme_booktabs() %>%
    fontsize(size = 9, part = "all") %>%
    align(align = "left", part = "all") %>%
    padding(padding = 4, part = "all")
  
  # Return both as a list for separate rendering
  list(Coefficients = ft_coef, Stats = ft_stats)
}



```

# STAT448 Assignment 2 – Regression Modelling Across Domains

This report addresses the three components of STAT448 Assignment 2, each focused on modelling using a different dataset. The project involves building, evaluating, and interpreting regression models using tools such as linear regression, stepwise selection, and regularised regression.

```{r child = 'doc/01_residential_modelling.Rmd'}
```

#```{r child = 'doc/02_parkinsons_modelling.Rmd'}
#```

#```{r child = 'doc/03_weather_modelling.Rmd'}
#```

## Conclusion

This report presented three focused modelling tasks corresponding to different datasets and modelling requirements. Each section includes analysis, model fitting, validation, and interpretation appropriate to the data and modelling technique.

Please refer to each child document for details on preprocessing, model evaluation, and final conclusions.